---
title: "{{DASHBOARD_TITLE}}"
format:
  dashboard:
    embed-resources: true
    theme: flatly
    scrolling: true
    engine: knitr
    expandable: false
---

:::{.hidden}
```{=html}
<link
  rel="stylesheet"
  href="https://cdn.jsdelivr.net/npm/bootstrap-icons@1.10.5/font/bootstrap-icons.css"
/>
```
:::

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(visNetwork)
library(tidyverse)
library(htmltools)
library(bsplus)
library(colorspace)
library(tippy)
library(reactable)
library(DT)
library(plotly)
library(igraph)
library(gprofiler2)

res_report      <- readRDS("{{PATH_RES_REPORT}}")
res_diffusion   <- readRDS("{{PATH_RES_DIFFUSION}}")
complete_network <- readRDS("{{PATH_COMPLETE_NETWORK}}")
se_to_severity  <- readRDS("{{PATH_SE_TO_SEVERITY}}")

# Package data
data("module_ppi_M1_20231211", package = "CENetwork")

# Node type colour palette
protein_color <- "#ffa500"
gene_color    <- "#ADD8E6"
go_color      <- "#D3D3D3"
pathway_color <- "#2f9ec2"
se_color      <- "#a8c71e"
drug_color    <- "#ff00b4"

apply_graph_custom <- function(g) {
  h <- g
  igraph::vertex_attr(h) <- igraph::vertex_attr(h) %>%
    as.data.frame() %>%
    mutate(color = case_when(
      type == "protein"      ~ protein_color,
      type == "gene"         ~ gene_color,
      type == "GO"           ~ go_color,
      type == "pathway"      ~ pathway_color,
      type == "side_effect"  ~ se_color,
      type == "drug/compound" ~ drug_color,
      TRUE ~ "grey"
    )) %>%
    mutate(prefix_type = case_when(
      type == "protein"      ~ "p:",
      type == "gene"         ~ "g:",
      type == "GO"           ~ "go:",
      type == "pathway"      ~ "r:",
      type == "side_effect"  ~ "se:",
      type == "drug/compound" ~ "d:",
      TRUE ~ ""
    )) %>%
    mutate(border = ifelse(TG | TF, "red", "grey")) %>%
    mutate(shape  = ifelse(input_diffusion, "star", "dot")) %>%
    mutate(frame.width = 5) %>%
    mutate(oldname = name) %>%
    mutate(name = paste0(prefix_type, display_name)) %>%
    as.list()
  return(h)
}
```

```{r prep, echo = FALSE, cache = FALSE}
g         <- res_diffusion$network
g_global  <- apply_graph_custom(g)
g_attr    <- igraph::vertex_attr(g) %>% as.data.frame() %>%
  left_join(se_to_severity, by = c("name" = "se"))

g_types    <- igraph::vertex_attr(g) %>% as.data.frame() %>%
  split(.$type) %>% imap(~.x %>% pull(name))

g_long_df <- igraph::as_long_data_frame(g)

# Subgraphs
ids_drugs_se  <- g_long_df %>% filter(from_type == "drug/compound") %>%
  { unique(c(.$from_name, .$to_name)) }
ids_drugs_tgt <- g_long_df %>% filter(to_type == "drug/compound") %>%
  { unique(c(.$from_name, .$to_name)) }
ids_pathway   <- g_long_df %>% filter(to_type == "pathway") %>%
  { unique(c(.$from_name, .$to_name)) }
ids_go_gene   <- g_long_df %>% filter(to_type == "GO", from_type == "gene") %>%
  { unique(c(.$from_name, .$to_name)) }
ids_go_prot   <- g_long_df %>% filter(to_type == "GO", from_type == "protein") %>%
  { unique(c(.$from_name, .$to_name)) }

g_drugs   <- igraph::induced_subgraph(g, c(ids_drugs_se, ids_drugs_tgt)) %>% apply_graph_custom()
g_pathway <- igraph::induced_subgraph(g, ids_pathway) %>% apply_graph_custom()
g_gos     <- igraph::induced_subgraph(g, c(ids_go_gene, ids_go_prot)) %>% apply_graph_custom()

g_proteins <- igraph::induced_subgraph(g, g_types$protein) %>% apply_graph_custom()
vertex_attr(g_proteins) <- vertex_attr(g_proteins) %>% as.data.frame() %>%
  left_join(module_ppi_M1_20231211, by = c("oldname" = "molecule")) %>%
  mutate(module = str_remove(module, ".*_") %>% as.numeric()) %>%
  mutate(color = module) %>%
  as.list()

# Input table
tab_input <- left_join(
  res_report$degree_signature,
  g_global %>% vertex_attr() %>% as.data.frame() %>% dplyr::select(oldname, type),
  by = c("name" = "oldname")
)

# Module enrichment table (graph level)
graph_module_info <- module_ppi_M1_20231211 %>%
  mutate(in_graph = molecule %in% g_types$protein) %>%
  group_by(module) %>%
  summarise(k = sum(in_graph), n = n()) %>%
  filter(k != 0) %>%
  arrange(desc(k)) %>%
  mutate(K = sum(k), N = sum(n)) %>%
  nest(count = -module) %>%
  mutate(contingency = imap(count, ~CENetwork:::make_contingency(k = .x$k, K = .x$K, n = .x$n, N = .x$N))) %>%
  mutate(pvalue = map_dbl(contingency, ~fisher.test(.x, alternative = "greater")$p.value)) %>%
  mutate(padj   = p.adjust(.$pvalue, method = "fdr")) %>%
  dplyr::select(module, count, pvalue, padj) %>%
  unnest(cols = count) %>%
  arrange(padj) %>%
  mutate(pvalue = round(pvalue, 2), padj = round(padj, 2))

# DILI / SE severity enrichment
va_complete <- complete_network %>% vertex_attr() %>% as.data.frame() %>%
  left_join(se_to_severity, by = c("name" = "se"))

stat_enrich_drugs <- tribble(
  ~name, ~ki, ~Ki, ~ni, ~Ni,
  "Ambiguous DILI",
    filter(g_attr,    vDILIConcern == "Ambiguous DILI-Concern") %>% pull(name),
    filter(va_complete, vDILIConcern == "Ambiguous DILI-Concern") %>% pull(name),
    g_attr$name, va_complete$name,
  "Less DILI",
    filter(g_attr,    vDILIConcern == "vLess-DILI-Concern") %>% pull(name),
    filter(va_complete, vDILIConcern == "vLess-DILI-Concern") %>% pull(name),
    g_attr$name, va_complete$name,
  "Most DILI",
    filter(g_attr,    vDILIConcern == "vMost-DILI-Concern") %>% pull(name),
    filter(va_complete, vDILIConcern == "vMost-DILI-Concern") %>% pull(name),
    g_attr$name, va_complete$name,
  "SE Severity: Mild",
    filter(g_attr,    max_severity == "Mild") %>% pull(name),
    filter(va_complete, max_severity == "Mild") %>% pull(name),
    g_attr$name, va_complete$name,
  "SE Severity: Moderate",
    filter(g_attr,    max_severity == "Moderate") %>% pull(name),
    filter(va_complete, max_severity == "Moderate") %>% pull(name),
    g_attr$name, va_complete$name,
  "SE Severity: Severe",
    filter(g_attr,    max_severity == "Severe") %>% pull(name),
    filter(va_complete, max_severity == "Severe") %>% pull(name),
    g_attr$name, va_complete$name,
  "SE Severity: Lifethreatening",
    filter(g_attr,    max_severity == "Lifethreatening") %>% pull(name),
    filter(va_complete, max_severity == "Lifethreatening") %>% pull(name),
    g_attr$name, va_complete$name,
  "SE Severity: Death",
    filter(g_attr,    max_severity == "Death") %>% pull(name),
    filter(va_complete, max_severity == "Death") %>% pull(name),
    g_attr$name, va_complete$name
) %>%
  mutate(
    k = map_dbl(ki, ~length(.x)),
    n = map_dbl(ni, ~length(.x)),
    K = map_dbl(Ki, ~length(.x)),
    N = map_dbl(Ni, ~length(.x))
  ) %>%
  group_by(name) %>%
  nest(count = c(k, K, n, N)) %>%
  mutate(contingency = imap(count, ~CENetwork:::make_contingency(k = .x$k, K = .x$K, n = .x$n, N = .x$N))) %>%
  mutate(p.value = map_dbl(contingency, ~fisher.test(.x, alternative = "greater")$p.value)) %>%
  mutate(padj    = p.adjust(p.value, method = "fdr")) %>%
  mutate(
    ki_length = imap_dbl(ki, ~length(.x)),
    ni_length = imap_dbl(ni, ~length(.x)),
    Ki_length = imap_dbl(Ki, ~length(.x)),
    Ni_length = imap_dbl(Ni, ~length(.x))
  ) %>%
  dplyr::select(ki_length, ni_length, Ki_length, Ni_length, p.value, padj)

# gprofiler enrichments
gost_res_go   <- gost(query = g_types$gene, organism = "hsapiens", sources = "GO",
                      highlight = TRUE, significant = FALSE)
gost_plot_go  <- gostplot(gost_res_go)

gost_res_path  <- gost(query = g_types$gene, organism = "hsapiens",
                       sources = c("KEGG", "REAC", "WP"),
                       highlight = TRUE, significant = FALSE)
gost_plot_path <- gostplot(gost_res_path)

# Permutation enrichment — joined into pathway/GO tables when available
# getFromNamespace works even if the function is not yet exported (dev workflow).
# CENetwork namespace is already loaded via data(..., package = "CENetwork") above.
.cde_fn <- tryCatch(
  getFromNamespace("compute_diffusion_enrichment", "CENetwork"),
  error = function(e) NULL
)

enrich_permut <- NULL
if (!is.null(.cde_fn) &&
    !is.null(res_diffusion$permutation_results) &&
    !is.null(res_diffusion$observed_visit_counts)) {
  enrich_permut <- tryCatch(
    .cde_fn(res_diffusion,
            target_types    = c("GO", "pathway"),
            min_visit_count = 1L),
    error = function(e) NULL
  )
}

permut_cols <- c("visit_count", "mean_perm", "sd_perm", "z_score", "pvalue", "padj")

pathways_display <- res_report$pathways
GOs_display      <- res_report$GOs

if (!is.null(enrich_permut)) {
  enrich_path <- enrich_permut %>% filter(type == "pathway") %>%
    dplyr::select(term, all_of(permut_cols))
  enrich_go   <- enrich_permut %>% filter(type == "GO") %>%
    dplyr::select(term, all_of(permut_cols))

  pathways_display <- pathways_display %>%
    left_join(enrich_path, by = c("name" = "term"))
  GOs_display <- GOs_display %>%
    left_join(enrich_go, by = c("name" = "term"))
}
```

```{r helpers, echo = FALSE}
with_tooltip <- function(value, tooltip, ...) {
  div(
    style = "text-decoration: underline; text-decoration-style: dotted; cursor: help",
    tippy::tippy(value, tooltip, ...)
  )
}

download_btn <- function(element_id) {
  tags$button(
    "Download as CSV",
    onclick = sprintf("Reactable.downloadDataCSV('%s')", element_id),
    style = "margin-top:10px; display:inline-block; width:auto; padding:6px 12px;
             font-size:14px; border-radius:4px; background-color:lightgrey;
             color:white; border:none; cursor:pointer;"
  )
}

row_color_style <- function(df) {
  function(index) {
    t <- df[index, "type"]
    col <- switch(t,
      "drug/compound" = drug_color,
      "gene"          = gene_color,
      "protein"       = protein_color,
      "GO"            = go_color,
      "side_effect"   = se_color,
      "pathway"       = pathway_color,
      NULL
    )
    if (!is.null(col)) list(background = colorspace::lighten(col, amount = 0.9))
  }
}

scrollable_reactable <- function(df, element_id, row_style_fn = NULL,
                                  fixed_row_style = NULL) {
  args <- list(
    data       = df,
    filterable = TRUE,
    wrap       = FALSE,
    resizable  = TRUE,
    searchable = FALSE,
    compact    = TRUE,
    pagination = FALSE,
    elementId  = element_id,
    fullWidth  = TRUE,
    height     = 1000,
    width      = "auto",
    onClick    = "expand",
    defaultColDef = reactable::colDef(maxWidth = 10000, minWidth = 100)
  )
  if (!is.null(row_style_fn))    args$rowStyle <- row_style_fn
  if (!is.null(fixed_row_style)) args$rowStyle <- fixed_row_style

  htmltools::tagList(
    tags$div(style = "height:400px; overflow-y:auto;",
             do.call(reactable::reactable, args)),
    tags$div(download_btn(element_id))
  )
}
```

# General informations

## Row {height=50%}

### Column {width=86%}

```{r}
#| title: "Diffusion network"
#| expandable: true
set.seed(123)
visNetwork::visIgraph(g_global, physics = FALSE) %>%
  visNetwork::visOptions(
    highlightNearest = list(enabled = TRUE, hover = FALSE),
    selectedBy = "type"
  )
```

### Column {width=14%}

### ValueBox Grid {layout-ncol=2}

```{r}
#| content: valuebox
#| title: "Number of nodes"
igraph::vcount(g)
```

```{r}
#| content: valuebox
#| title: "Input nodes"
list(
  value = sum(g_attr$input_diffusion),
  title = htmltools::HTML('<i class="bi bi-diagram-3"></i> Input nodes')
)
```

```{r}
#| content: valuebox
#| title: "Nb. Genes"
list(color = gene_color, value = sum(g_attr$type == "gene"))
```

```{r}
#| content: valuebox
#| title: "Nb. Proteins"
list(color = protein_color, value = sum(g_attr$type == "protein"))
```

### Column {width=7%}

```{r}
#| content: valuebox
#| title: "Nb. Pathways"
list(color = pathway_color, value = sum(g_attr$type == "pathway"))
```

```{r}
#| content: valuebox
#| title: "Nb. Drugs"
list(color = drug_color, value = sum(g_attr$type == "drug/compound"))
```

```{r}
#| content: valuebox
#| title: "Nb. GOs"
list(color = go_color, value = sum(g_attr$type == "GO"))
```

```{r}
#| content: valuebox
#| title: "Nb. Side Effects"
list(color = se_color, value = sum(g_attr$type == "side_effect"))
```

## Row {height=50%}

### Column {width=30%}

```{r}
#| title: "General enrichment"
#| expandable: true
try(res_report$radar_plots_data$interest %>%
      CENetwork:::make_ggradar(max.upper.limit = 30, interactive = TRUE))
```

### Column {width=70% .tabset .flow}

#### Input

<div style="height: 450px; overflow-y: auto;">

```{r}
#| results: asis
scrollable_reactable(
  tab_input %>% dplyr::select(display_name, name, everything()),
  element_id    = "degree-table",
  row_style_fn  = row_color_style(tab_input),
  fixed_row_style = NULL
)
```

</div>

#### All vertex attributes

<div style="height: 450px; overflow-y: auto;">

```{r}
#| results: asis
scrollable_reactable(
  g_attr,
  element_id   = "attr-table",
  row_style_fn = row_color_style(g_attr)
)
```

</div>

#### Most visited nodes

<div style="height: 450px; overflow-y: auto;">

Each input seed generates a path to the final subnetwork; some nodes may be visited multiple times.

```{r}
#| results: asis
scrollable_reactable(
  res_report$visite_autoraute,
  element_id   = "attr-autoroute",
  row_style_fn = row_color_style(res_report$visite_autoraute)
)
```

</div>

# Drugs

## Row {height=50%}

### Column {width=50%}

```{r}
#| title: "Drugs — Targets | Drugs — Side-effects"
set.seed(123)
visNetwork::visIgraph(g_drugs, physics = FALSE) %>%
  visNetwork::visOptions(highlightNearest = list(enabled = TRUE, hover = FALSE))
```

### Column {width=50%}

```{r}
#| title: "DILI and SE Severity enrichment"
DILI_enrich_plot <- ggplot(
  stat_enrich_drugs %>%
    mutate(type = ifelse(str_detect(name, "SE"), "SE Severity", "DILI")) %>%
    mutate(shape = ifelse(ki_length == 0, "blue", "red")) %>%
    mutate(name = str_remove(name, "SE Severity: ")) %>%
    mutate(name = factor(name, levels = rev(.$name))),
  aes(y = name, x = -log10(padj))
) +
  geom_point(aes(color = shape)) +
  ylab("") +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "grey") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_color_identity() +
  facet_wrap(type ~ ., dir = "v", scales = "free_y") +
  coord_cartesian(xlim = c(0, -log10(0.00001)))

ggplotly(DILI_enrich_plot)
```

## Row {height=50%}

### Column {width=100% .tabset .flow}

#### Drug info

<div style="height: 450px; overflow-y: auto;">

```{r}
#| results: asis
scrollable_reactable(
  res_report$drugs,
  element_id      = "attr-drugs-infos",
  fixed_row_style = list(background = colorspace::lighten(drug_color, amount = 0.9))
)
```

</div>

#### Targets

<div style="height: 450px; overflow-y: auto;">

```{r}
#| results: asis
scrollable_reactable(
  res_report$drugs.targets,
  element_id      = "attr-drugs-targets",
  fixed_row_style = list(background = colorspace::lighten(protein_color, amount = 0.9))
)
```

</div>

#### Side Effects

<div style="height: 450px; overflow-y: auto;">

```{r}
#| results: asis
scrollable_reactable(
  res_report$drugs.side_effect,
  element_id      = "attr-drugs-SE",
  fixed_row_style = list(background = colorspace::lighten(se_color, amount = 0.9))
)
```

</div>

#### Shortest distance from input

<div style="height: 450px; overflow-y: auto;">

```{r}
#| results: asis
scrollable_reactable(
  res_report$drug.dist.signature %>% arrange(distance),
  element_id = "attr-drugs-dist-input"
)
```

</div>

# Pathways

## Row {height=50%}

```{r}
#| title: "Pathways — Proteins"
set.seed(123)
visNetwork::visIgraph(g_pathway, physics = FALSE) %>%
  visNetwork::visOptions(
    highlightNearest = list(enabled = TRUE, hover = FALSE),
    nodesIdSelection = TRUE
  )
```

```{r}
#| title: "Pathway global enrichment"
gost_plot_path
```

## Row {height=50%}

### Column {width=100% .tabset .flow}

#### Pathway info

<div style="height: 450px; overflow-y: auto;">

```{r}
#| results: asis
scrollable_reactable(
  pathways_display,
  element_id      = "attr-pathway",
  fixed_row_style = list(background = colorspace::lighten(pathway_color, amount = 0.9))
)
```

</div>

#### Shortest distance from input

<div style="height: 450px; overflow-y: auto;">

```{r}
#| results: asis
scrollable_reactable(
  res_report$pathway.dist.signature,
  element_id = "attr-pathway-dist-signature"
)
```

</div>

# GO terms

## Row {height=50%}

```{r}
#| title: "GO — Genes | GO — Proteins"
set.seed(123)
visNetwork::visIgraph(g_gos, physics = FALSE) %>%
  visNetwork::visOptions(highlightNearest = list(enabled = TRUE, hover = FALSE))
```

```{r}
#| title: "GO global enrichment"
gost_plot_go
```

## Row {height=50%}

### Column {width=100% .tabset .flow}

#### GO terms info

<div style="height: 450px; overflow-y: auto;">

```{r}
#| results: asis
scrollable_reactable(
  GOs_display,
  element_id      = "attr-go",
  fixed_row_style = list(background = colorspace::lighten(go_color, amount = 0.9))
)
```

</div>

#### Shortest distance from input

<div style="height: 450px; overflow-y: auto;">

```{r}
#| results: asis
scrollable_reactable(
  res_report$GO.dist.signature,
  element_id = "attr-go-dist-signature"
)
```

</div>

# Modules

## Row {height=50%}

```{r}
#| title: "Modules sub-network"
#| expandable: true
set.seed(123)
visNetwork::visIgraph(g_proteins, physics = FALSE) %>%
  visNetwork::visOptions(
    highlightNearest = list(enabled = TRUE, hover = FALSE),
    nodesIdSelection = TRUE,
    selectedBy = "module"
  )
```

```{r}
#| title: "Module Overrepresentation"
Module_enrich_plot <- ggplot(
  graph_module_info %>%
    mutate(shape = ifelse(k == 0, "blue", "red")) %>%
    arrange(desc(padj)) %>%
    mutate(module = factor(module, levels = rev(.$module))),
  aes(y = module, x = -log10(padj))
) +
  geom_point(aes(color = shape)) +
  ylab("") +
  geom_vline(xintercept = -log10(0.05), linetype = "dotted", color = "grey") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_color_identity() +
  coord_cartesian(xlim = c(0, -log10(0.00001)))

ggplotly(Module_enrich_plot)
```

## Row {height=50%}

```{r}
#| title: "Enrichment in Toxic modules"
try(res_report$radar_plots_data$module_GOTox %>%
      CENetwork:::make_ggradar(max.upper.limit = 30, interactive = TRUE))
```

```{r}
#| title: "Module enrichment table"
DT::datatable(graph_module_info)
```
